Solar Flares Across the Spectrum: Observations and Instrumentation

### 2025-04-11 introduction/intro.tex
Local changes

The Sun is our nearest astrophysical object, which serves as a natural laboratory for various branches of Physics - atomic physics, spectroscopy, turbulence, magneto-hydrodynamics, stellar evolution, dynamo and magnetism and plays a key role in space and terrestrial weather. The Sun is also the primary source of light and energy on earth. So it is only natural that we try to understand the different physical processes involved in the energectics of the Sun, which indirectly affects life on Earth. In this chapter I briefly introduce the Sun and various layers of its atmosphere. At the end we provide a motivation for the thesis.

The Sun is a G2V star, with surface luminosity $3.86~\times~10^{26}$ W and an effective temperature of $\sim 5780~K$. The Sun is mostly made out of Hydrogen(92.1\%) and Helium(7.8\%) along with negligible quantities of heavier elements C, N, O, Mg, Si, Ne and Fe. The study of the Sun can be mainly divided into three components: the Solar interior, the Solar atmosphere and the Heliosphere.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The solar interior}\label{solar_int}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the innermost zone of the star the Sun has a temperature of about 15 MK and a density of $1.6~\times~10^{5}~km.m^{-3}$ that fuses Hydrogen into Helium mainly by p-p cycle \citep{bethe38} and to some extent by CNO cycle \citep{bethe39}.

%% ############# %%
\begin{figure}[h!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_int.png}
    \caption{A schematic depiction of the different layers of the Sun. The figure also shows some of the most prominent magnetic structures on the Sun e.g sunspot, prominence, coronal hole. Credit : NASA/Goddard.}
    \label{fig_solar_int}
\end{figure}
%% ############# %%

Fig.~\ref{fig_solar_int} shows a schematic diagram of various layers of the Sun. Beyond the core we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized. The high energy $\gamma$-ray photons generated in the core, have very small mean free path in this region, as they collide numerous times eventually thermalizing becoming visible photon. Beyond the radiative zone the degree of ionization of Hydrogen and Helium changes drastically giving rise to a sharp temperature gradient. This is the convective zone where the energy from the inner layers of the Sun is transported adiabetically to the surface of the Sun. In between the radiative and the convective zone, there is a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tacholine, which is believed to play a major role in generating Sun's magnetic field via the Dynamo process \citep{corbard01}.

The core and the radiative zone of the Sun rotates as a solid body, whereas the convective zone rotates differentially. The rotation period varies from $\sim~25~days$ to $\sim~35~days$ from the equator to the pole.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The solar atmosphere}\label{solar_atmos}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The layers outside the convection zone together, form the solar atmosphere - the Photosphere, which is also known as the Sun's surface, the Chromosphere, the transition region and the Corona. While we can define geometrical height or depths to define these layers, it is much more useful to define them using the local optical depths of various spectral lines, which are directly used to observe the layers. This also gives us a idea about which spectral features originate mainly from which portions of the solar atmosphere. Fig.~\ref{fig_solar_atm} shows the variation of temperature and number density across various layers of the Solar atmosphere \sr{Ref. here}.

%% ############# %%
\begin{figure}[h!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_atm.png}
    \caption{The variation of electron temperature (solid line) and electron density ($\mamthrm{N_{e}}$, dashed line) and density of neutral hydrogen atoms ($\mathrm{N_{H}}$, dotted- dashed line) in the solar atmosphere as derived based on the 1-D model calculations by \sr{Ref. here}.}
    \label{fig_solar_atm}
\end{figure}
%% ############# %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Photosphere}\label{photosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Photosphere is defined to be the layer where $\tau_{\lambda~\sim~5000}~=~1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400-600 km deep and has an effective temperature of 5780 K. The magnetic filed lines arising from the tacholine penetrate the Photosphere and creates a `carpet' over the whole region \citep{priest14}. As mentioned earlier, this is the Solar surface known as the `Quite Sun'(QS). This region exhibits an average magnetic flux density of \textbf{10 -- 50~G}. The QS surface is covered with cells of roughly four size- granule, meso granules, super granules and giant cells. Some regions exhibit much stronger magnetic flux density often associated with highly twisted magnetic field structures. These magnetic features are manifested as sunspots, spicules, bright points etc.

As the density changes drastically at the Photosphere compared to the convective zone, the thermalized photons from Sun's core start free streaming again, as the mean free path also increases drastically and as a result perturbs the thermal equilibrium (TE from hereon). So, we have to define the thermodynamic quantities of the Photosphere in Local Thermal Equilibrium (LTE from hereon). As we move away from the Photosphere, because the density keeps on decreasing steadily, the LTE conditions also start deviating similarly \citep{philips08}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Chromosphere}\label{chromosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Chromosphere is a highly non-uniform dynamic layer with a thickness of 1500~--~3000~km, with increasing temperature (upto $10^{4}$ K) and decreasing number density. As seen in the fig. \ref{fig_solar_atm}, the temperature of the Chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by creation of shock waves \citep{carlsson07}. The Chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, dominance of wave motions \sr{Ref. here}.

%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={1cm 3cm 2cm 5cm},clip,width=0.8\textwidth]{Figures/sji_images.pdf}
    \caption{IRIS SJI Slit-jaw images of NOAA AR 13521 on Dec 21st, 2023 in \ion{Mg}{2} 2796 {\AA}~(left panel) showing the upper Chromosphere. The right panel shows the same active region in 2832 {\AA}~continuum corresponding to upper photosphere.}
    \label{fig:sji_features}
\end{figure}
%%%%%%%%%%%%%%%%

In fig.~\ref{fig:sji_features} we show the NOAA AR 13521 in IRIS SJI 2796 {\AA} (left panel) and 2832 {\AA} (right panel). The dark sunspots and plage regions surrounding it are clearly visible in the 2832 {\AA} wavelength window. In the 2796 {\AA} window we see the plage along with light bridges more clearly.

@@

The transition region above the Chromosphere is a thin, dynamic layer where we witness a dramatic increase in temperature by two orders of magnitude along with a similar decrease in electron density by nearly 5-6 orders of magnitude which is demonstrated in fig. \ref{fig_solar_atm}. In general the this layer is roughly $\sim$ 100 km in thickness, but that can change depending on various dynamic conditions of Chromosphere($\sim~10^{4}~K$) below, or Corona($\sim~10^{6}~K$) on top. The transition region is generally characterized by the steep change in the temperature, pressure gradients, drastic change in local optical depth and competition between gas pressure and magnetic pressure. The transition region also manifests the ARs as various magnetic features such as small scale brightening, jets, spicules, fibrils etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Corona}\label{corona}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The outermost layer of the Solar atmosphere is the Corona. The Corona is only visible in naked eyes during the total solar eclipse or via Coronagraphs. It maintains a steady temperature of 1-2 Mk during ambient conditions, but can rise upto tens of MK during solar flares. The coronal continuum spectrum is produced due to Thompson scattering of photons from photosphere off coronal electrons. Corona also emits in various emission lines, e.g. Mg, Fe, Si, S, Ca, O etc. across a wide band of electromagnetic spectrum ranging from X-ray to visible wavelengths.

Similar to the other layers, the corona exhibits a variety of structures such as ARs, coronal holes, bright points, filaments, flares. The flare arcades are most prominently visible in the Corona. So, most of our existing studies over the last decade has focused on the coronal features of the flares. In fig.~\ref{fig:sji_features} we show one of the most well known and well studied solar eruptions from Aug 31st, 2012. A giant prominence erupted from the south east limb. The erupted prominence material was visible in all AIA channels from 304 \AA~(Chromospheric He II, see fig.~\ref{fig:aia-feature} \textbf{panel (a)}) across all six Coronal channels (\textbf{panel (b){--}(g)}). The sun spots were also clearly visible in HMI magnetogram and continuum (see fig.~\ref{fig:aia-feature} \textbf{panel (h)}). In \textbf{panel (i)} we show the co-aligned AIA 304 \AA, 131 \AA and HMI magnetogram image. The sunspots lie at the base of the erupted prominence, accompanied by the hot thermal plasma associated with the accompanying flare. The novelty of such observation lies in the clear spatial connection across various layers of the sun, connected via energy and momentum transported from one layer to the other.

%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{Figures/aia_features.pdf}
    \caption{Image of the prominence eruption and associated flare from NOAA AR 11562 on Aug 31st, 2012$\sim$19:50 UT. The prominence is visible in all of the AIA channels. We see the coronal holes in AIA 335 \AA~\textbf{(panel c)}. The flare loops of the associated event is visible in AIA 94 \AA~\textbf{(panel e)} and AIA 131 \AA~\textbf{(panel f)}. The sunspots are visible in HMI continuum \textbf{panel (h)}. We show a co aligned map of AIA 304 \AA, HMI magnetogram and AIA 131 \AA~in \textbf{panel (i)}. The spatial association of the sunspots with the hot flareloop plasma and the prominence eruption is clearly demostrated here. }
    \label{fig:aia-feature}
\end{figure}
%%%%%%%%%%%%%

@@

Solar flares are the most powerful magnetic events in the solar system. They are described as a sudden increase in brightness in localized areas on Sun. Within tens of minutes, they can release over $10^{32}$ erg of energy, which is emitted across the entire electromagnetic spectrum from radio to gamma rays. They can also launch high energetic particles into the interplanetary medium. Most of the flares occur in magnetic active regions, and the amount of flare energy released is comparable to the free energy stored in the magnetic system. The term "flare" is generally used explicitly for the entire magnetically-driven event's electromagnetic radiation, as it is the most significant fraction of the total energy liberated. The total energy released varies from event to event. It is also known that larger events occur much less frequently than smaller events.

%% ##################################################### %%
\subsection{Brief history of flare observation}\label{sol_flr_1}
%% ##################################################### %%

In September 1, 1859, R.C. Carrington and R. Hodgson observed the first flare in the continuum of white light\citep{carrington1859,hodgson1859}. The localized brightening on the Sun have remained an enigma ever since. We have been observing the local flaring events across all wavelengths from both ground and space based observatories.  Shortly after the observation by Carrington and Hodgson, people started studying the Sun extensively in the H$\alpha$ line which is formed in the Chromosphere, and the reports of flaring events became more and more frequent and progressively more and more complex. No two events were similar, as there were variations observed in source of size, ejections of plasma along with shock waves driving into the interplanetary space. Advances of radio technology during the second world war ensured detection of presence of non-thermal electrons in the solar corona, during military radar operations \citep{hey46}. Around the same time S.E. Forbrush noticed ground level cosmic ray enhancement during major solar flares. These discoveries eluded that the flaring events do not only involve the thermal plasma, but is somehow connects with high energy particles and involves the corona. in the 1950s we started observing the Sun in hard X-rays ($\ge~10~keV$) with rockets and balloons. \cite{peterson59} discovered the first hard X-ray emission during a flare in 1958. Later on, it was deduced from the observations of the enhancements observed in radio and hard X-rays that, the the ejected energetic particles may contain a substantial fraction of the initial energy released \citep{brown71}. The hard X-ray is created by the bremsstrahlung of the electrons colliding into dense material, resulting in a power-law energy distribution. The broadband radio emission from 1 to 100 GHz is created from gyrosynchrotron emission. Finally, observations in EUV and soft X-ray($\le$~10~keV) have shown that the energy released form the flare heats the plasma contained in the coronal loops to temperatures beyond 30 MK.

%% ################################################################# %%
\subsection{Neupert effect}\label{npt_eff}
%% ################################################################# %%

\cite{neupert68} observed a curious correlation that, the soft X-ray flux during the rise phase of the flare is proportional to the time integral of the centimeter radio flux since the start of the flare. The centimeter radio flux is emitted by relativistic electrons. So, later on the same correlation was found between the hard X-ray flux and soft X-ray flux and can be expressed as,

@@

This empirical relation is known as the ``Neupert effect''. \cite{neupert68} had already suggested that this may be due to a causal relationship between the thermal plasma and the energetic electrons. The logical explanation is: the soft X-ray mainly originates from a thermal plasma heated by the energy of the flaring event deposited by the flare accelerated electrons. It is important to mention that we know now, Eqn.\ref{npt_eff_eq} is only valid if cooling by conduction of radiation is negligible.

%% ################################################################# %%
\subsection{``Standard'' model of solar flare}\label{sol_flr_std_mod}
%% ################################################################# %%

There were several observations like Neupert effect, which would help us to constrain several phenomenon observed in the flares:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.5\linewidth]{Figures/phu_63_8_818_f2.jpg}
    \caption{A schematic diagram of the standard model of solar flare (adapted from \cite{lysenko20}).}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%
\begin{enumerate}
    \item The magnetic reconnetion happens in corona, releasing the magnetic free energy. Electrons and energetic particles from the reconnetion site are accelerated along the realigned magnetic field lines. The accelerated particles that escape along the open field lines towards the earth gives rise to the particle events seen from earth.
    \item The accelerated particles move along the field lines downwards along the magnetic field lines, giving rise to the microwave observations seen from flares via gyro magnetic radiation.
    \item The accelerated particles hit the Chromosphere which is considerably denser than the Corona, giving rise to Hard X-ray and gamma ray observed from the foot points. This also explains why the coronal hard X-ray source is considerably softer than the foot points. The energetic particles deposit their energy into the local Chromosphere, as they go through series of collision and eventually thermalize.
    \item  As the plasma thermalizes, it starts emitting in thermal bremsstrahlung giving rise to the soft X-ray observed. Also, as the energy is deposited into the Chromosphere from the accelerated particles, it heats the local Chromosphere environment it gradually increases the local pressure. 
    \item As the pressure grows, when the pressure gradient builds up enough the local plasma starts expanding upwards(essentially due to buoyancy) and slowly fills up the coronal loops with soft X-ray emitting plasma. This phenomenon is known as "Chromospheric evaporation". This was directly observed later on, in blue-shifted lines of hot material.
    \item This whole scenario explains the Neupert effect. As the energy form the accelerated particles is converted into the soft X-ray emitting plasma, and it builds up over time. That explains the soft X-ray being proportional to the time integral of the hard X-ray flux.
\end{enumerate}
%%%%%%%%%%%%%%%%%

This whole scenario is known as the ``Standard'' model of solar flare. While the standard model is an attempt to explain and unify various kind of differences seen from the numerous flares we observe, there are several cases where the standard model cannot explain the observations. For example, there have been observations of flares where the hard X-ray foot point do not form. It is proposed that in these cases the plasma in the coronal loops is so dense that the accelerated particle collide enough to thermalize within the flare loops before reaching the Chromosphere, defusing the energy more evenly within the loop, rather than dumping it at the base of the loops \citep{veronig02,veronig04}. Another flare which previously occurred at the same region might explain the denser coronal loops \citep{strong84,bone07}.

%% ################################################################# %%
\subsection{The energectics of solar flares}\label{sol_flr_energ}
%% ################################################################# %%

>>>>

The Sun is our nearest astrophysical object, which serves as a natural laboratory for various branches of Physics - atomic physics, spectroscopy, turbulence, magneto-hydrodynamics, stellar evolution, dynamo and magnetism and plays a key role in space and terrestrial weather. The Sun is also the primary source of light and energy on Earth. So it is only natural that we try to understand the different physical processes involved in the energetics of the Sun, which directly affects life on Earth. In this chapter, I briefly introduce the Sun, various layers of its atmosphere and Solar flares. Finally, we provide a motivation for the thesis and a brief outline.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Sun}\label{Sun}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Sun is a G2V star, with surface luminosity $3.86~\times~10^{26}$~W and an effective temperature of $\sim 5780~K$. The Sun is mostly made out of Hydrogen (92.1\%) and Helium (7.8\%), along with negligible quantities of heavier elements C, N, O, Mg, Si, Ne and Fe, among others. The study of the Sun can be divided into three components, \textit{viz.} the Solar interior, the Solar atmosphere and the Heliosphere. Fig.~\ref{fig_solar_int} shows a schematic diagram of various regions of the Sun. Below, we briefly describe these three regions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar interior}\label{solar_int} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\info{First you have to say what is the interior! For example the interior is made of three regions -- the core, the radiative zone and the convective zone. The innermost zone, i.e., the core}
Depending on various physical characteristics, the Solar interior can be divided into three zones {\it e.g.} the core, radiative zone and convective zone. In the innermost zone, the core, the Sun has a temperature of about 15~MK and a density of $\mathrm{1.6\times 10^{5}~km.m^{-3}}$. Here Hydrogen fuses into Helium primarily by p-p chain reaction \citep{bethe38} and to some extent by CNO cycle \citep{bethe39}. Beyond the core, we have the optically thick radiative zone, where the Hydrogen and Helium are completely ionized due to very high density. The high energy $\gamma$-ray photons generated in the core have a very small mean free path in this region, as they collide numerous times and propagate across this region. As we move radially outwards, temperature and density both decrease slowly due to gravitational equilibrium. Due to this reduction of temperature, there is a decrease in the thermal motion of the particles, and the hydrogen and helium start to recombine, giving rise to an increase in opacity at the top of the radiative zone. This change in opacity makes the propagation of photons via radiation less efficient, as photons get trapped and reabsorbed more. This creates a sharp temperature gradient due to the "photon pileup". This temperature gradient gives rise "convective instability". This is the convective zone where the energy from the inner layers of the Sun is transported adiabatically to the surface of the Sun. 
%In between the radiative and the convective zone, there is \sroy{believed to be }a thin layer, roughly at $\sim~0.7~R_{\odot}$ known as the tachocline. \sroy{This region} is believed to play a major role in generating Sun's magnetic field via the Dynamo process \citep{corbard01}.

%% ############# %%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={2cm 11cm 0.5cm 8cm}, clip, width=0.9\textwidth]{solar_int.pdf}
    \caption[A schematic depiction of the different regions of the Sun.]{A schematic depiction of the different regions of the Sun, including some of the most prominent magnetic structures on the Sun e.g., Sunspot, prominence, coronal hole.}
    \label{fig_solar_int}
\end{figure}
%% ############# %%

The core and the radiative zone of the Sun rotate as a solid body, whereas the convective zone rotates deferentially. The rotation period varies from $\sim~25~days$ to $\sim~35~days$ from the equator to the pole.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Solar atmosphere}\label{solar_atmos}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The layers outside the convection zone together form the Solar atmosphere and is further divided into different regions. The photosphere, which is also known as the Sun's surface, the chromosphere, the transition region and the corona. While we can define geometrical height or depths to define these layers, it is much more useful to define them using the local optical depths of various spectral lines, which are directly used to observe the layers. This also gives us an idea about which spectral features originate mainly from which portions of the Solar atmosphere. Fig.~\ref{fig_solar_atm} shows the variation of temperature and number density across various layers of the Solar atmosphere \citep{philips08}.

%% ############# %%
\begin{figure}[ht!]
    \centering
    \includegraphics[width = 0.8\linewidth]{Figures/solar_atm.png}
    \caption[Density and temperature profiles across various layers of the Sun.]{The variation of electron temperature (solid line) and electron density ($\mathrm{N_{e}}$, dashed line) and density of neutral hydrogen atoms ($\mathrm{N_{H}}$, dot-dashed line) in the Solar atmosphere \citep{philips08}.}
    \label{fig_solar_atm}
\end{figure}
%% ############# %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The photosphere}\label{photosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The photosphere is defined to be the layer where $\tau_{\lambda~\sim~5500}=1$. This is in the green part of the visible spectrum and the Sun is opaque in visible beyond this layer, hence it is called to be the surface of the Sun. This layer is 400{--}600~km thick and has an effective temperature of 5780 K. The magnetic field lines arising from the tachocline penetrate the photosphere and create a `carpet' over the whole region \citep{priest14}. As mentioned earlier, this is the Solar surface known as the `quiet Sun'(QS). This region exhibits an average magnetic flux density of 10 {--} 50 G. The QS surface is covered with cells of roughly four sizes, namely, granules, mesogranules, supergranules and giant cells.

Some regions exhibit much stronger magnetic flux density, often associated with highly twisted magnetic field structures. These magnetic features are manifested as Sunspots. The Sunspots harbour some of the strongest concentrated magnetic fields, reaching up to several kilogauss in strength. The other noteworthy Solar magnetic structure apart from these are believed to be spatially unresolved flux tubes composing a network with field strengths $\sim$ {\it $O(10^{3} G)$}\citep{solanki93,grossman96,rubio19}.

As the density changes drastically at the photosphere compared to the convective zone, the thermalized photons from the Sun's core start free streaming, as the mean free path also increases drastically and, as a result, perturbs the thermal equilibrium (TE). Therefore, the thermodynamic quantities of the photosphere are defined under the assumption of Local Thermal Equilibrium (LTE). As we move higher from the photosphere, the LTE conditions also start deviating because the density keeps on decreasing steadily \citep{philips08}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The chromosphere}\label{chromosphere}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The chromosphere is a highly non-uniform dynamic layer with a thickness of 1500 {--} 3000 km, with increasing temperature (up to $10^{4}$ K) and decreasing number density. As seen in Fig.~\ref{fig_solar_atm}, the temperature of the chromosphere saturates before the dramatic rise in the transition region. This saturation is attributed to a steady deposition of acoustic energy by the creation of shock waves \citep{carlsson07}. The chromosphere also exhibits a sharp gradient in the plasma $\beta$ factor, non-LTE conditions, and the dominance of wave motions. The observable features in the chromosphere are active regions, sunspots, plages and also light bridges as labelled in images taken by IRIS SJI (slit-jaw imager) 2796 {\AA} (left panel) and 2832 {\AA} (right panel) displayed in Fig.~\ref{fig:sji_features}.

%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[trim={1cm 3cm 2cm 5cm},clip,width=0.95\textwidth]{Figures/sji_images.pdf}
    \caption[IRIS SJI observation of NOAA AR 13521 on Dec 21st, 2023.]{IRIS slit-jaw images of NOAA AR 13521 taken in \ion{Mg}{2} 2796 {\AA} (left panel) and 2832 {\AA}~continuum (right panel) showing the upper chromosphere and upper photosphere, respectively. Different features are labelled.} 
    \label{fig:sji_features}
\end{figure}
%%%%%%%%%%%%%%%%

@@

The transition region above the chromosphere is a thin, dynamic layer where we witness a dramatic increase in temperature by two orders of magnitude along with a similar decrease in electron density, demonstrated in Fig.~\ref{fig_solar_atm}. In general, this layer is roughly $\sim$100~km in thickness, but that can change depending on various dynamic conditions of chromosphere ($\sim~10^{4}~K$) below, and corona ($\sim~10^{6}~K$) on top. The transition region is generally characterized by steep changes in temperature, pressure gradients, local optical depth, and competition between gas and magnetic pressure. The transition region also manifests the ARs as various magnetic features such as small-scale brightening, jets, spicules, fibrils etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The corona}\label{corona}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The outermost layer of the atmosphere is the corona. The corona maintains a temperature of 1-2~MK during ambient conditions but can rise up to tens of MK during explosive eruptions such as Solar flares. Due to such high temperatures, the coronal gas is fully ionized. It is only visible to the naked eye during the total Solar eclipse or with coronagraphs due to Thomson scattering of the photosphere with free electrons in the corona. corona also emits in various emission lines, e.g. Mg, Fe, Si, S, Ca, O, etc., across a wide band of electromagnetic spectrum ranging from X-ray to visible wavelengths. Similar to the other layers, the corona exhibits a variety of structures such as ARs, coronal holes, bright points, filaments, and flares. The flare arcades are most prominently visible in the corona. %\info{I made a lot of changes in this section. please see if all is consistent.}

%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{Figures/aia_features.pdf}
    \caption[Different parts of a flare eruption observed across various wavelengths]{AIA observations of prominence eruption and associated flare as observed on Aug 31st, 2012$\sim$19:50 UT. Different features across different wavelengths have been labelled.}
    %\caption[Different parts of a flare eruption observed across various wavelengths.]{Image of the prominence eruption and associated flare from NOAA AR 11562 on Aug 31st, 2012$\sim$19:50 UT. The prominence is visible in all of the AIA channels. We see the coronal holes in AIA 335 \AA~\textbf{(panel c)}. The flare loops of the associated event is visible in AIA 94 \AA~\textbf{(panel e)} and AIA 131 \AA~\textbf{(panel f)}. The Sunspots are visible in HMI continuum \textbf{panel (h)}. We show a co aligned map of AIA 304 \AA, HMI magnetogram and AIA 131 \AA~in \textbf{panel (i)}. The spatial association of the Sunspots with the hot flareloop plasma and the prominence eruption is clearly demonstrated here. }
    \label{fig:aia-feature}
\end{figure}
%%%%%%%%%%%%%

%\info{I would write the caption of Fig.~\ref{fig:aia-feature} as: AIA observations of prominence eruption and associated flare as observed on Aug 31st, 2012$\sim$19:50 UT. Different features have been labelled.}

@@

Solar flares are the most powerful energetic explosions in the Solar system. They are described as a sudden increase in brightness in localized areas of the Sun. Within tens of minutes, they can release over $10^{32}$ erg of energy, which is emitted across the entire electromagnetic spectrum, i.e., from radio to gamma rays. They can also launch highly energetic particles into the interplanetary medium. Flares occur in magnetic active regions. It has been estimated that the amount of flare energy released is comparable to the free energy stored in the magnetic system \citep{emslie12,ash17}. %The term "flare" is generally used explicitly for the entire magnetically-driven event's electromagnetic radiation, as it is the most significant fraction of the total energy liberated. 
The total energy released varies from event to event. It is also known that larger events occur much less frequently than smaller events.

Fig.~\ref{fig:sji_features} displays one of the most well-known and well-studied Solar eruptions from Aug 31st, 2012. During this event, an associated giant prominence erupted from the southeast limb, which was visible in all AIA channels as shown in Fig.~\ref{fig:aia-feature}. The sunspots were also clearly visible and labelled in the HMI continuum (see Fig.~\ref{fig:aia-feature}h). In Fig.~\ref{fig:aia-feature}i, we show the co-aligned AIA 304 {\AA} and HMI continuum observation. The Sunspots lie at the base of the erupted prominence, accompanied by the hot thermal plasma associated with the accompanying flare (visible in Fig.~\ref{fig:aia-feature}e {--} f). The novelty of such observation lies in the clear spatial connection across various layers of the Sun, connected via energy and momentum transported from one layer to the other.

%------------------------------------
\subsection{Brief history of flare observation}\label{sol_flr_1}
%------------------------------------

On September 1, 1859, R.C. Carrington and R. Hodgson independently observed the first flare in white-light continuum~\citep{carrington1859,hodgson1859}. Such localized brightening on the Sun has remained an enigma ever since. 
Shortly after the observation by Carrington and Hodgson, the Sun was being studied extensively in the H$\alpha$ wavelength, which essentially images the chromosphere, and the reports of flaring events became increasingly frequent and progressively more complex. No two events were similar, as there were variations observed in the size of the source, ejections of plasma, and shock waves driving into the interplanetary space. Advances in radio technology during the Second World War affirmed the detection of the presence of non-thermal electrons in the Solar corona during military radar operations \citep{hey46}. Around the same time, S.E. Forbrush noticed ground-level cosmic ray enhancement during major Solar flares. These discoveries eluded that the flaring events do not only involve the thermal plasma but somehow connect with high energy particles and involve the corona.

In the late 1950s, using rockets and balloons, the Sun was observed with hard X-rays ($\ge~10~keV$). \cite{peterson59} discovered the first hard X-ray emission during a flare in 1958. Later on, it was deduced from the enhancements observed in radio and hard X-rays that the ejected energetic particles may contain a substantial fraction of the initial energy released \citep{brown71}. Note that while the hard X-ray is created by the bremsstrahlung radiation of the electrons colliding into dense material, resulting in a power-law energy distribution, the broadband radio emission from 1 to 100 GHz is created from gyrosynchrotron emission. Flares are now routinely observed with different space-borne instruments, in particular in the EUV and soft X-ray($\le$~10~keV), which shows that the energy released from the flare heats the local plasma to temperatures beyond 30~MK.

%% ################################################################# %%
%\subsection{Neupert effect}\label{npt_eff} \info{I would not give it a name of new subsection}
%% ################################################################# %%

\cite{neupert68} observed a curious correlation that the soft X-ray flux during the rising phase of the flare is proportional to the time integral of the centimetre radio flux since the start of the flare, suggesting a correlation between thermal plasma and the energetic electrons. Note that the centimetre radio flux is emitted by relativistic electrons. A similar correlation was found between the hard X-ray and soft X-ray flux later on. Such correlation can be expressed as,

@@

\noindent and is known as the ``Neupert effect''. This relationship shows that the soft X-ray radiation primarily originates from a thermal plasma which is heated by the energy of the flaring event deposited by the accelerated electrons. We note, however, that Eqn.\ref{npt_eff_eq} is only valid if cooling by conduction is negligible and/or the role of ions in the flare energy deposition is minimal \citep{veronig05}.

%% ################################################################# %%
\subsection{``Standard'' model of Solar flare}\label{sol_flr_std_mod}
%% ################################################################# %%

Based on various multi-wavelength studies of Solar flares, combined with theoretical modelling a `standard' model of Solar flares has been advocated by \cite{carmichael64,sturrock66,hirayama74,kopp76}, and termed as \textsl{CSHKP} model. The flares occur due to the abrupt release of magnetic free energy, which was previously stored in the coronal magnetic field via flux emergence and surface motions \citep{forbes06}. The majority of the strongest flares are eruptive in nature. The `standard' model attributes the release of flare energy to magnetic reconnection in the corona, which happens due to the eruption and/or coronal mass ejection \citep{shibata95,lin_n_forbes00,moore01,priest_n_forbes02}. It is noteworthy that, despite continuous attempts over the decades, this is not at all a quantitative model. Rather, it is an attempt to unify the multi-wavelength observations and modelling attempts in an orderly manner. Naturally, the model can not explain all the observations from various flares. A schematic picture of the model is shown in Fig.~\ref{fig:std_mod}. Below we describe the salient points of the `standard' model:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[ht!]
    \centering
    \hspace{5cm}
    \includegraphics[trim={1cm 8cm 0cm 8cm}, clip, width=0.95\linewidth]{Figures/std_mod.pdf}
    \caption{A schematic diagram of the standard model of Solar flare.}
    \label{fig:std_mod}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%

%\info{This whole description of Standard Model needs to be rewritten. You are going too lateral. Here, the aim is not to bring out the caveats, etc, but to outline the standard model. As you say that you are giving the salient features... so just provide that! Not more than 2-3 sentences in each bullet.}

%\begin{enumerate}
    %\item In the standard model, following the reconnection event occurring that releases the magnetic free energy, electrons and ions from the reconnection site are accelerated along the realigned magnetic field lines. The accelerated particles that escape along the open \info{field lines do not open during flares! Please read Aly and Sturocck paper on this! Very interesting concept!} field lines towards the Earth give rise to the particle events seen from Earth \info{which particles are you talking about? There are different kinds of particles -- particles accelerated due to shock waves that come very fast ... remember proton showers on the detector and other particles which are frozen in the field... they come later}. The coronal reconnection source where the energy is released is often observed in HXR. The loop-top HXR \info{are these same as the reconnection source?} sources are often accompanied by the thermal SXR sources. The loop top HXR source is observed above the coronal SXR source \citep{masuda94}. Later on, most of the HXR sources were reported to be cospatial with the SXR source \citep{bataglia_n_benz07}. This thermal source is distinctly different from the post-flare SXR and EUV loops. This thermal coronal source appears before the HXR emission begins, completely contrary to the "Neupert" behaviour. This SXR loop top source is a key behaviour of the pre-flare phase. The emission measure of these sources is higher than typical coronal levels, suggesting evaporation from the chromosphere before the HXR emission. This also suggests that during the early phases of a flare, the energy is conducted gradually by thermal particles, compared to the energetic non-thermal particles later on.
